---
title: "STEP 4. Statistical Analysis"
author: "Veronika Juhasz"
date: "2/5/2022"
output: html_document
---

### Loading packages and checking directory

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("tidyverse", "dplyr", "ggplot2", "lme4", "lattice", "ggrepel")
getwd()
```


### Reading in data and shaping it up for later work ###

```{r}
SV_df_raw <- read_csv("./data/NOW_preprocessed/SV_BLM_bias_connotation_combined.csv") %>% 
  rename("Perspective_ws" = "Perspective(ws)") %>% 
  select(ID, Date, S_string, V_lemmatized, Perspective_ws, media_bias_rating, Side)

VO_df_raw <- read_csv("./data/NOW_preprocessed/VO_BLM_bias_connotation_combined.csv") %>% 
  rename("Perspective_wo" = "Perspective(wo)") %>% 
  select(ID, Date, O_string, V_lemmatized, Perspective_wo, media_bias_rating, Side)

print(sum(is.na(SV_df_raw$media_bias_rating)))
print(sum(is.na(VO_df_raw$media_bias_rating)))


# SV datasets

SV_df <-  SV_df_raw %>% 
  drop_na(media_bias_rating) %>% 
  filter(Side != "0") %>% 
  mutate(
    SO_BLM_role = Side) %>% 
  rename(
    Subject_or_object = S_string,
    Empathy_score = Perspective_ws,
    Political_alignment_rating = media_bias_rating)
SV_df$SO_BLM_role[SV_df$SO_BLM_role == 1] <- "anti-BLM-SO"
SV_df$SO_BLM_role[SV_df$SO_BLM_role == -1] <- "pro-BLM-SO"

# VO datasets

VO_df <-  VO_df_raw %>% 
  drop_na(media_bias_rating) %>% 
  filter(Side != "0") %>% 
  mutate(
    SO_BLM_role = Side) %>% 
  rename(
    Subject_or_object = O_string,
    Empathy_score = Perspective_wo,
    Political_alignment_rating = media_bias_rating)
VO_df$SO_BLM_role[VO_df$SO_BLM_role == 1] <- "anti-BLM-SO"
VO_df$SO_BLM_role[VO_df$SO_BLM_role == -1] <- "pro-BLM-SO"

# combining into one full dataset

total_df <- rbind(SV_df, VO_df)

proBLM_SO_df <- total_df %>% 
  filter(SO_BLM_role == "pro-BLM-SO")
antiBLM_SO_df <- total_df %>% 
  filter(SO_BLM_role == "anti-BLM-SO")

# changing the order of the levels of political alignment

SV_df$Political_alignment_rating <- factor(SV_df$Political_alignment_rating, levels = c("Left", "Lean Left", "Center", "Lean Right", "Right"))
VO_df$Political_alignment_rating <- factor(VO_df$Political_alignment_rating, levels = c("Left", "Lean Left", "Center", "Lean Right", "Right"))
total_df$Political_alignment_rating <- factor(total_df$Political_alignment_rating, levels = c("Left", "Lean Left", "Center", "Lean Right", "Right"))
proBLM_SO_df$Political_alignment_rating <- factor(proBLM_SO_df$Political_alignment_rating, levels = c("Left", "Lean Left", "Center", "Lean Right", "Right"))
antiBLM_SO_df$Political_alignment_rating <- factor(antiBLM_SO_df$Political_alignment_rating, levels = c("Left", "Lean Left", "Center", "Lean Right", "Right"))
```

### Doing some data cleaning and reshaping the data for plotting ###

```{r}

total_df_filtered <- total_df %>% 
  dplyr::filter(grepl("^Breonna$|^Breonna Taylor$|^Floyd$|^George Floyd$|^movement$|^Black Lives Matter$|^BLM$|^protester$|^protesters$|^demonstrator$|^demonstrators$|^police$|^National Guard$|^Blue Lives Matter$|^government$" , Subject_or_object, ignore.case = TRUE)) %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Breonna$", "Breonna Taylor") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Floyd$", "George Floyd") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Trayvon$", "Trayvon Martin") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^movement$", "Black Lives Matter") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Movement$", "Black Lives Matter") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^BLM$", "Black Lives Matter") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^blm$", "Black Lives Matter") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^protesters$", "protester") %>%
  mutate_at("Subject_or_object", str_replace_all, "^Protesters$", "protester") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Protester$", "protester") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^PROTESTERS$", "protester") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^demonstrators$", "demonstrator") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Demonstrators$", "demonstrator") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Demonstrator$", "demonstrator") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Police$", "police") %>% 
  mutate_at("Subject_or_object", str_replace_all, "^Government$", "government")


table(total_df_filtered$Subject_or_object)

total_df_filtered_forplot <- total_df_filtered %>% 
  mutate(
      Political_alignment_of_news_source = Political_alignment_rating) %>% 
  filter(
    Political_alignment_of_news_source != "Center") %>% 
  mutate_at("Political_alignment_of_news_source", str_replace_all, "^Left$", "Liberal") %>% 
  mutate_at("Political_alignment_of_news_source", str_replace_all, "^Lean Left$", "Liberal") %>% 
  mutate_at("Political_alignment_of_news_source", str_replace_all, "^Right$", "Conservative") %>% 
  mutate_at("Political_alignment_of_news_source", str_replace_all, "^Lean Right", "Conservative")

table(total_df_filtered_forplot$Political_alignment_of_news_source)

total_df_filtered_forplot_liberal <-  total_df_filtered_forplot %>%
  filter(Political_alignment_of_news_source == "Liberal") %>% 
  group_by(Subject_or_object) %>% 
  summarise(Empathy_score = mean(Empathy_score))  %>% 
  rename(Empathy_score_liberal = Empathy_score)
total_df_filtered_forplot_liberal$SO_BLM_role <- ifelse(
  grepl("police|National Guard|government", total_df_filtered_forplot_liberal$Subject_or_object), "anti-BLM-SO", "pro-BLM-SO")

total_df_filtered_forplot_conservative <-  total_df_filtered_forplot %>%
  filter(Political_alignment_of_news_source == "Conservative") %>% 
  group_by(Subject_or_object) %>% 
  summarise(Empathy_score = mean(Empathy_score)) %>% 
  rename(Empathy_score_conservative = Empathy_score)
total_df_filtered_forplot_conservative$SO_BLM_role <- ifelse(
  grepl("police|National Guard|government", total_df_filtered_forplot_conservative$Subject_or_object), "anti-BLM-SO", "pro-BLM-SO")

summary_df <- left_join(total_df_filtered_forplot_liberal, total_df_filtered_forplot_conservative, by = "Subject_or_object")
```

### Plotting ###

```{r}
interaction.plot(x.factor = total_df_filtered$Political_alignment_rating, #x-axis variable
                 trace.factor = total_df_filtered$SO_BLM_role, #variable for lines
                 response = total_df_filtered$Empathy_score, #y-axis variable
                 fun = mean, #metric to plot
                 ylab = "Mean empathy score",
                 xlab = "Political Alignment Rating of news source",
                 col = c("red", "blue"),
                 lty = 1, #line type
                 lwd = 2, #line width
                 trace.label = "Pro- or anti BLM figures")


ggplot(total_df_filtered, aes(x= Political_alignment_rating, y = Empathy_score, group = SO_BLM_role, color = SO_BLM_role)) +
  stat_summary(fun.y = mean,
               geom = "point",
               width = 0.6) +
  stat_summary(fun.y = mean,
               geom = "line",
               alpha = 0.5) +
  stat_summary(fun.data = mean_se,
               geom = "errorbar",
               width = 0.2) +
  scale_color_manual(name = "BLM-related figures' role\nin the BLM movement",
                         labels=c("pro-BLM-SO" = "pro-BLM role",
                                  "anti-BLM-SO" = "anti-BLM role"),
                         values = c("pro-BLM-SO" = "blue", 
                                    "anti-BLM-SO" = "red")) +
  labs(x = "\nPolitical alignment rating of news source", y = "Mean empathy score towards pro- and anti-BLM figures\n", title = "Mean Empathy Score Towards BLM-related Figures in News Media with \nDifferent Political Alignment - with SE Error Bars" ) +
  theme_minimal()


 
ggplot(summary_df, aes(x = Empathy_score_conservative, y = Empathy_score_liberal, color = SO_BLM_role.y, label = Subject_or_object)) +
  geom_point() +
  geom_abline() +
  geom_label_repel(aes(label = Subject_or_object),
                  box.padding   = 0.3, 
                  point.padding = 0.3,
                  segment.color = 'grey50') +
  theme(legend.position = "left") +
#  expand_limits(y = c(0.5, 0.5)) +
  scale_color_manual(name = "BLM-related figures' role\nin the BLM movement",
                         labels=c("pro-BLM-SO" = "pro-BLM role",
                                  "anti-BLM-SO" = "anti-BLM role"),
                         values = c("pro-BLM-SO" = "blue", 
                                    "anti-BLM-SO" = "red")) +
  labs(x = "\nMean empathy score by conservative news sources", y = "Mean empathy score by liberal news sources\n", title = "Mean Empathy Score Towards BLM-related Figures in News Media" ) +
  theme_minimal()


```
### Making the model ###

```{r}
m1 <- lmerTest::lmer(Empathy_score ~ Political_alignment_rating * SO_BLM_role + (1|ID), data = total_df_filtered)
summary(m1)

round(summary(m1)$coefficients, 5)
```

## A more interpretable version of the same model, with political alignment being a continuous varibale instead of categorical ###

```{r}
total_df_filtered_cont <- total_df_filtered %>%
  mutate(Political_alignment_rating_cont = as.numeric(Political_alignment_rating))
  
m1_cont <- lmerTest::lmer(Empathy_score ~ Political_alignment_rating_cont * SO_BLM_role + (1|ID), data = total_df_filtered_cont)
summary(m1_cont)

round(summary(m1_cont)$coefficients, 5)
```

# Main effect plot

```{r}

total_df_filtered_forplot_main <-  total_df_filtered_forplot
total_df_filtered_forplot_main$SO_BLM_role <- ifelse(
  grepl("police|National Guard|government", total_df_filtered_forplot_main$Subject_or_object), "anti-BLM role", "pro-BLM role")

ggplot(total_df_filtered_forplot_main, aes(x = SO_BLM_role, y = Empathy_score, color = SO_BLM_role)) +
  stat_summary(fun.y = mean,
               geom = "point",
               width = 0.6)  +
  stat_summary(fun.data = mean_se,
               geom = "errorbar",
               width = 0.2) +
  scale_color_manual(values = c("red", "blue")) +
  labs(x = "\nBLM-related figures' role in the BLM movement", y = "Mean empathy score in news reporting\n", title = "Main Effect of BLM-related Figures Pro- or Anti-BLM Role on\nEmpathy Score with SE Error Bars" ) +
  theme_minimal()

```

### Creating plot for each pro- and anti-BLM figure's empathy score

```{r}
total_df_filtered_forplot_cons <- total_df_filtered_forplot %>% 
  filter(SO_BLM_role == "anti-BLM-SO")

total_df_filtered_forplot_cons_sum <- total_df_filtered_forplot_cons %>% # the names of the new data frame and the data frame to be summarised
  group_by(Political_alignment_of_news_source, Subject_or_object) %>% # the grouping variable
  dplyr::summarise(
    mean = mean(Empathy_score), # calculates the mean of each group
    sd = sd(Empathy_score), # calculates the standard deviation of each group
    n_es = n(), # calculates the sample size per group
    SE = sd(Empathy_score) / sqrt(n())
  ) # calculates the standard error of each group

  
# with white error bar

ggplot(total_df_filtered_forplot_cons_sum, aes(x = Subject_or_object, y = mean, fill = Political_alignment_of_news_source)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = mean - SE, ymax = mean + SE), width = 0.5, size = 0.5, position = position_dodge(0.7)) +
  geom_errorbar(aes(ymin = mean - SE, ymax = mean), width = 0.0, size = 0.5, position = position_dodge(0.7), color = "white") +
  theme(legend.position = "left") +
  scale_fill_manual(name = "Political Alignment \nof News Outlet",
                         values = c("Liberal" = "blue", 
                                    "Conservative" = "red")) +
  labs(x = "\nBLM-related figures (anti-BLM)", y = "Mean empathy score in  news reporting\n", title = "Mean Empathy Score Towards Anti-BLM Figures", subtitle = "with SE Error Bars" ) +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", size = "10", angle= 25, hjust=1, color = "black"))


total_df_filtered_forplot_lib <- total_df_filtered_forplot %>% 
  filter(SO_BLM_role == "pro-BLM-SO")

total_df_filtered_forplot_lib_sum <- total_df_filtered_forplot_lib %>% # the names of the new data frame and the data frame to be summarised
  group_by(Political_alignment_of_news_source, Subject_or_object) %>% # the grouping variable
  dplyr::summarise(
    mean = mean(Empathy_score), # calculates the mean of each group
    sd = sd(Empathy_score), # calculates the standard deviation of each group
    n_es = n(), # calculates the sample size per group
    SE = sd(Empathy_score) / sqrt(n())
  ) # calculates the standard error of each group


# with white error bar

ggplot(total_df_filtered_forplot_lib_sum, aes(x = Subject_or_object, y = mean, fill = Political_alignment_of_news_source)) +
  geom_bar(position = "dodge", stat = "identity", width = 0.7) +
  geom_errorbar(aes(ymin = mean - SE, ymax = mean + SE), width = 0.5, size = 0.5, position = position_dodge(0.7)) +
  geom_errorbar(aes(ymin = mean - SE, ymax = mean), width = 0.0, size = 0.5, position = position_dodge(0.7), color = "white") +
  theme(legend.position = "left") +
  scale_fill_manual(name = "Political Alignment \nof News Outlet",
                         values = c("Liberal" = "blue", 
                                    "Conservative" = "red")) +
  labs(x = "\nBLM-related figures (pro-BLM)", y = "Mean empathy score in news reporting\n", title = "Mean Empathy Score Towards Pro-BLM Figures", subtitle = "with SE Error Bars" ) +
  theme_minimal() +
  theme(axis.text.x = element_text(face = "bold", size = "10", angle= 25, hjust=1, color = "black"))

  
```

### Checking current version of R and how to cite it ###

```{r}
citation()
R.version
```


